---
title: "Attendance Report"
subtitle: "`r params$report_type_label`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: "Actserv Analytics"

params:
  report_type:  "weekly_flash"
  date_from:    ""
  date_to:      ""
  dept_scope:   "All departments"
  data_json:    "[]"

format:
  pdf:
    documentclass: article
    geometry:
      - top=20mm
      - left=20mm
      - right=20mm
      - bottom=20mm
    fontsize: 10pt
    colorlinks: true
    linkcolor: "f89734"
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{longtable}
        \usepackage{array}
        \usepackage{xcolor}
        \usepackage{colortbl}
        \usepackage{fancyhdr}
        \usepackage{graphicx}
        \usepackage{float}
        \usepackage{caption}
        \definecolor{navy}{HTML}{244a4a}
        \definecolor{brandblue}{HTML}{f89734}
        \definecolor{present}{HTML}{007A63}
        \definecolor{late}{HTML}{C47D00}
        \definecolor{absent}{HTML}{C0303A}
        \definecolor{lightgray}{HTML}{EEF3F7}
        \definecolor{bordergray}{HTML}{E0E5E9}
        \pagestyle{fancy}
        \fancyhf{}
        \renewcommand{\headrulewidth}{0.4pt}
        \fancyhead[L]{\textcolor{navy}{\textbf{Actserv} Attendance Analytics}}
        \fancyhead[R]{\textcolor{gray}{\small CONFIDENTIAL}}
        \fancyfoot[L]{\textcolor{gray}{\small \leftmark}}
        \fancyfoot[R]{\textcolor{gray}{\small Page \thepage}}

execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(data.table)
library(knitr)
library(kableExtra)
library(lubridate)
library(ggplot2)

# ── Load data from JSON param ─────────────────────────────────────────────────
df <- tryCatch(
  data.table::setDT(jsonlite::fromJSON(params$data_json)),
  error = function(e) data.table()
)

# ── Report type label ─────────────────────────────────────────────────────────
type_labels <- list(
  weekly_flash = "Weekly Attendance Flash",
  monthly_performance = "Monthly Performance Report",
  dept_summary = "Department Summary",
  employee_detail = "Employee Attendance Detail",
  late_log = "Late Arrivals Log",
  absent_log = "Absent Days Log"
)
report_label <- type_labels[[params$report_type]] %||% params$report_type

# date range labels:
date_from_fmt <- if (nchar(params$date_from) > 0) {
  format(as.Date(params$date_from), "%d %b %Y")
} else {
  "—"
}
date_to_fmt <- if (nchar(params$date_to) > 0) {
  format(as.Date(params$date_to), "%d %b %Y")
} else {
  "—"
}

# helpers:
`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

pct <- function(x) paste0(round(x, 1), "%")
hrs <- function(x) paste0(round(x, 1), "h")

color_pct <- function(x, low = 70, mid = 85) {
  dplyr::case_when(
    x < low ~ "\\textcolor{absent}{",
    x < mid ~ "\\textcolor{late}{",
    TRUE ~ "\\textcolor{present}{"
  )
}
```

\begin{center}
{\color{navy}\rule{\linewidth}{2pt}}
\vspace{4pt}
{\color{navy}\Large\textbf{`r report_label`}}\\[4pt]
{\color{gray}\normalsize `r date_from_fmt` — `r date_to_fmt`}\\[2pt]
{\color{gray}\small Scope: `r params$dept_scope`}
\vspace{4pt}
{\color{navy}\rule{\linewidth}{0.5pt}}
\end{center}

\vspace{6pt}

```{r report-dispatch}
#| results: asis

# Route to the appropriate section based on report_type
report_type <- params$report_type
```

```{r kpi-summary}
#| results: asis

# KPI summary banner — shown for all report types except detail logs
if (
  report_type %in%
    c("weekly_flash", "monthly_performance", "dept_summary", "employee_detail")
) {
  if (nrow(df) > 0 && "on_schedule" %in% names(df)) {
    adherence <- round(mean(df$on_schedule, na.rm = TRUE) * 100, 1)
    avg_hours <- round(mean(df$hours_in_office, na.rm = TRUE), 1)
    late_count <- if ("status" %in% names(df)) {
      sum(df$status == "late", na.rm = TRUE)
    } else {
      0L
    }
    absent_ct <- if ("status" %in% names(df)) {
      sum(df$status == "absent", na.rm = TRUE)
    } else {
      0L
    }
    employees <- if ("userid" %in% names(df)) {
      data.table::uniqueN(df$userid)
    } else {
      0L
    }

    cat(sprintf(
      "
\\vspace{4pt}
\\noindent
\\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
\\hline
\\rowcolor{lightgray}
\\textbf{\\small Employees} & \\textbf{\\small Adherence} & \\textbf{\\small Avg Hrs/Day} & \\textbf{\\small Late Arrivals} & \\textbf{\\small Absent Days} \\\\
\\hline
\\centering\\large\\textbf{%d} & \\centering\\large\\textbf{%s} & \\centering\\large\\textbf{%s} & \\centering\\large\\textbf{%d} & \\centering\\large\\textbf{%d} \\\\
\\hline
\\end{tabular}
\\vspace{8pt}
",
      employees,
      pct(adherence),
      hrs(avg_hours),
      late_count,
      absent_ct
    ))
  }
}
```

---

```{r section-weekly-flash}
#| results: asis

if (report_type == "weekly_flash") {
  cat("## Weekly Attendance Summary\n\n")

  if (nrow(df) == 0) {
    cat("*No data available for the selected period.*\n")
  } else {
    # Daily summary
    if ("date" %in% names(df) && "status" %in% names(df)) {
      df[, date := as.Date(date)]

      daily_summary <- df[,
        .(
          present = sum(status == "present", na.rm = TRUE),
          late = sum(status == "late", na.rm = TRUE),
          absent = sum(status == "absent", na.rm = TRUE),
          wfh = sum(status == "wfh", na.rm = TRUE),
          half_day = sum(status == "half_day", na.rm = TRUE),
          adherence = round(mean(on_schedule, na.rm = TRUE) * 100, 1)
        ),
        by = date
      ][order(date)]

      daily_summary[, date := format(date, "%a %d %b")]

      kbl(
        daily_summary,
        col.names = c(
          "Date",
          "Present",
          "Late",
          "Absent",
          "WFH",
          "Half Day",
          "Adherence %"
        ),
        booktabs = TRUE,
        longtable = TRUE,
        align = "lcccccc"
      ) |>
        kable_styling(
          latex_options = c("hold_position", "repeat_header"),
          font_size = 9
        ) |>
        column_spec(
          7,
          color = ifelse(
            daily_summary$adherence < 70,
            "C0303A",
            ifelse(daily_summary$adherence < 85, "C47D00", "007A63")
          )
        ) |>
        print()
    }

    # Top 5 late arrivals
    if ("status" %in% names(df)) {
      late_df <- df[status == "late"][order(-minutes_late)][1:min(5, .N)]
      if (nrow(late_df) > 0) {
        cat("\n\n### Top Late Arrivals This Week\n\n")
        late_display <- late_df[, .(
          Employee = employee_name,
          Department = department,
          Date = format(as.Date(date), "%a %d %b"),
          Arrived = arrival_time,
          `Mins Late` = minutes_late
        )]
        kbl(late_display, booktabs = TRUE, align = "llccc") |>
          kable_styling(latex_options = "hold_position", font_size = 9) |>
          print()
      }
    }
  }
}
```

```{r section-monthly}
#| results: asis

if (report_type == "monthly_performance") {
  cat("## Monthly Performance Report\n\n")

  if (nrow(df) == 0) {
    cat("*No data available for the selected period.*\n")
  } else {
    # Department summary
    if ("department" %in% names(df)) {
      cat("### Department Overview\n\n")

      dept_tbl <- df[,
        .(
          Staff = data.table::uniqueN(userid),
          `Adh %` = round(mean(on_schedule, na.rm = TRUE) * 100, 1),
          `Avg Hrs` = round(mean(hours_in_office, na.rm = TRUE), 1),
          Late = sum(status == "late", na.rm = TRUE),
          Absent = sum(status == "absent", na.rm = TRUE)
        ),
        by = .(Department = department)
      ][order(-`Adh %`)]

      kbl(dept_tbl, booktabs = TRUE, align = "lccccc") |>
        kable_styling(latex_options = "hold_position", font_size = 9) |>
        column_spec(
          3,
          color = ifelse(
            dept_tbl$`Adh %` < 70,
            "C0303A",
            ifelse(dept_tbl$`Adh %` < 85, "C47D00", "007A63")
          )
        ) |>
        print()
    }

    # Monthly trend
    if ("date" %in% names(df)) {
      df[, week := isoweek(as.Date(date))]
      weekly <- df[,
        .(
          adherence = round(mean(on_schedule, na.rm = TRUE) * 100, 1),
          late = sum(status == "late", na.rm = TRUE),
          absent = sum(status == "absent", na.rm = TRUE)
        ),
        by = week
      ][order(week)]

      cat("\n\n### Weekly Trend\n\n")
      kbl(
        weekly,
        col.names = c("Week", "Adherence %", "Late Count", "Absent Count"),
        booktabs = TRUE,
        align = "cccc"
      ) |>
        kable_styling(latex_options = "hold_position", font_size = 9) |>
        print()
    }
  }
}
```

```{r section-dept-summary}
#| results: asis

if (report_type == "dept_summary") {
  cat("## Department Summary\n\n")

  if (nrow(df) == 0) {
    cat("*No data available for the selected period.*\n")
  } else {
    display_cols <- intersect(
      c(
        "department",
        "headcount",
        "schedule_adherence",
        "avg_hours_in_office",
        "late_arrivals",
        "absent_days",
        "wfh_days",
        "attendance_rate"
      ),
      names(df)
    )
    tbl <- df[, ..display_cols][order(-schedule_adherence)]
    setnames(
      tbl,
      display_cols,
      c(
        "Department",
        "Staff",
        "Adherence %",
        "Avg Hrs",
        "Late",
        "Absent",
        "WFH",
        "Attend %"
      )[seq_along(display_cols)]
    )
    kbl(tbl, booktabs = TRUE, longtable = TRUE, align = "lcccccc") |>
      kable_styling(
        latex_options = c("hold_position", "repeat_header"),
        font_size = 9
      ) |>
      print()
  }
}
```

```{r section-employee-detail}
#| results: asis

if (report_type == "employee_detail") {
  cat("## Employee Attendance Detail\n\n")

  if (nrow(df) == 0) {
    cat("*No data available for the selected period.*\n")
  } else {
    display_cols <- intersect(
      c(
        "employee_name",
        "department",
        "days_worked",
        "schedule_adherence",
        "avg_arrival_time",
        "avg_leaving_time",
        "avg_hours_in_office",
        "late_count",
        "absent_days"
      ),
      names(df)
    )
    tbl <- df[, ..display_cols][order(department, -schedule_adherence)]
    setnames(
      tbl,
      display_cols,
      c(
        "Employee",
        "Department",
        "Days",
        "Adh %",
        "Arrival",
        "Leaving",
        "Avg Hrs",
        "Late",
        "Absent"
      )[seq_along(display_cols)]
    )
    kbl(tbl, booktabs = TRUE, longtable = TRUE, align = "llcccccc") |>
      kable_styling(
        latex_options = c("hold_position", "repeat_header", "striped"),
        font_size = 8.5
      ) |>
      print()
  }
}
```

```{r section-late-log}
#| results: asis

if (report_type == "late_log") {
  cat("## Late Arrivals Log\n\n")
  cat(sprintf("*Total late arrivals: %d*\n\n", nrow(df)))

  if (nrow(df) == 0) {
    cat("*No late arrivals in the selected period.*\n")
  } else {
    display_cols <- intersect(
      c("employee_name", "department", "date", "arrival_time", "minutes_late"),
      names(df)
    )
    tbl <- df[, ..display_cols][order(date, -minutes_late)]
    tbl[, date := format(as.Date(date), "%a %d %b %Y")]
    setnames(
      tbl,
      display_cols,
      c("Employee", "Department", "Date", "Arrived", "Mins Late")[seq_along(
        display_cols
      )]
    )
    kbl(tbl, booktabs = TRUE, longtable = TRUE, align = "llccc") |>
      kable_styling(
        latex_options = c("hold_position", "repeat_header", "striped"),
        font_size = 9
      ) |>
      column_spec(
        5,
        bold = TRUE,
        color = ifelse(tbl$`Mins Late` > 30, "C0303A", "C47D00")
      ) |>
      print()
  }
}
```

```{r section-absent-log}
#| results: asis

if (report_type == "absent_log") {
  cat("## Absent Days Log\n\n")
  cat(sprintf("*Total absent records: %d*\n\n", nrow(df)))

  if (nrow(df) == 0) {
    cat("*No absences in the selected period.*\n")
  } else {
    display_cols <- intersect(
      c("employee_name", "department", "date"),
      names(df)
    )
    tbl <- df[, ..display_cols][order(department, date)]
    tbl[, date := format(as.Date(date), "%a %d %b %Y")]
    setnames(
      tbl,
      display_cols,
      c("Employee", "Department", "Date")[seq_along(display_cols)]
    )
    kbl(tbl, booktabs = TRUE, longtable = TRUE, align = "llc") |>
      kable_styling(
        latex_options = c("hold_position", "repeat_header", "striped"),
        font_size = 9
      ) |>
      print()
  }
}
```

---

\vspace{6pt}
\begin{center}
{\color{gray}\footnotesize
Generated by Actserv Attendance Analytics \textbf{v`r app_version`} $\cdot$
`r format(Sys.time(), "%d %b %Y %H:%M")` $\cdot$
Data sourced from BioTime via Parquet data lake
}
\end{center}
